---
title: "Statistical Modelling"
author: "Kamau Kamau"
date: "22/10/2019"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,message =F,warning  =F)
```

```{r}
# library(MASS)
# head(oldfaithful)
# df$pred <- -54.461 + 12.016*df$rm
# 
# ggplot(df,mapping = aes(y=medv,x=rm)) + 
#   geom_point() +
#   geom_abline(intercept = -54.461, slope = 12.016, color = "red") +
#   geom_linerange(aes(ymin = df$medv, ymax = df$pred), color = "gray")
```

# Assumptions for the Regression Model

What are the assumptions that we have made using this model?

A straight line was the best way to model the overall relationship between log price and log carat.
The noise terms  εi  all have the same variance.
The noise terms  εi  are normally distributed.
The error terms are independent. (i.i.d. stands for "independent and identically distributed".)
We have names for each of the assumptions:

* Linearity
* Constant variance (homoscedasticity)
* Normality
* Independence

```{r}
library(tidyverse)
data(diamonds)
head(diamonds)
diamonds %>% ggplot(aes(carat,price))+geom_point()+scale_y_log10()+scale_x_log10()

diam_lm <-lm(log(price)~log(carat),data =diamonds)
summary(diam_lm)
plot(diam_lm,which =1) #line should be roughly flat ** test for linearlity
```

# Advanced Dummy variables

```{r}
library(tidyverse)
AdX_team  <- tribble(
  ~Person, ~Pet, 
  "Lewis", "Cat",
  "Sash", "Cat",
  "David", "Dog",
  "Jono", "Rabbit",
  "Dee", "Rabbit",
  "Alex", "Dog"
)
AdX_team  <- AdX_team %>% mutate(Pet = factor(Pet))
AdX_team
library(modelr)
model_matrix(AdX_team,~Pet)

```

# Additional Data
```{r}
data("chickwts")
chickwts  <- as_tibble(chickwts)
head(chickwts)
ggplot(chickwts,aes(feed, weight)) + geom_boxplot()+geom_jitter()

#fit a linear model
chickwts.lm  <- lm(weight ~ feed, data = chickwts)
summary(chickwts.lm)

# test if does feed have an effect on mean weight gain of chicks?
anova(chickwts.lm)
```

# predicting with mixed variables: both categorial as well as continous

```{r}
library(mlbench)
data("BostonHousing2")
BostonHousing2 <- as_tibble(BostonHousing2)
BostonHousing2
#Use just the number of rooms first
BostonHousing2.lm  <- lm(medv ~ rm, data = BostonHousing2)
summary(BostonHousing2.lm)

# add the Chas (Charles river factor) where its 1 if the house has a view to the river otherwise 0
BostonHousing2.lm2  <- lm(medv ~ rm + chas, data = BostonHousing2)
summary(BostonHousing2.lm2)

# does more rooms equal havng a view?
ggplot(BostonHousing2,mapping = aes(rm,medv)) + 
  geom_point(aes(color = chas)) + 
  geom_smooth(method = 'lm', mapping = aes(color=chas))
# check this more quantitavely using interaction terms
summary(lm(medv ~ rm + chas + rm:chas, data = BostonHousing2))

summary(lm(formula = price ~ carat * factor(cut, ordered = FALSE), data = diamonds))# with an interaction term
```

# stratified Datasets - the Gapminder problem
“How does life expectancy change over time for each country?”

```{r}
library(gapminder)
ggplot(gapminder, aes(year, lifeExp, group = country)) +
    geom_line(alpha = 0.25)
# one Country
au <- filter(gapminder, country == "Australia")
ggplot(au,aes(year, lifeExp)) + geom_line()
au.lm <- lm(lifeExp ~ year, data = au)
summary(au.lm)


# nest the dataframes such that each country has its own dataset subset
gapminder.nested  <- gapminder %>% 
  group_by(country, continent) %>% 
  nest()
gapminder.nested
# how to get the nested data
gapminder.nested$data[3]
# create a function for paralell model creation
country_model <- function(df) {
  lm(lifeExp ~ year, data = df)
}
gapminder.model  <- gapminder.nested %>% 
  mutate(model = map(data, country_model), 
         coef = map(model, broom::tidy))
gapminder.model$coef[[1]]
# unnest the dataframe
gapminder.coef  <- gapminder.model %>% unnest(coef)
gapminder.coef


```

```{r}
gapminder.coef %>% 
  mutate(term = fct_recode(term, 
                           Intercept = "(Intercept)", 
                           Slope = "year")) %>% 
  ggplot(aes(estimate, fill = term)) + 
  geom_density(show.legend = FALSE, alpha = 0.5) + 
  geom_histogram(col = "black", fill = "lightgrey",
                 alpha = 0.5,
                 aes(y = ..density..)) + 
  facet_wrap(~term, scales = "free") + 
  scale_fill_brewer(palette = "Set1") + 
  theme_minimal() + 
  labs(y = NULL, x = "Estimate")
```

